### 2024-02-09 doc/15yrsChange.Rmd
Add Recruitment Failure Paragraph

>>>>

Recruitment Failure Paragraph

White-tailed deer (Odocoileus virginianus) represent an endemic nuisance species prevalent in eastern deciduous forests. Their populations have substantially increased in the past 50 years (McShea et al. 1997), due in large part to human influence. (Found in McGarvey et al. 2013 – Brown et al. 2000, Côté et al. 2004, Rooney 2001).  White-tailed deer preferentially browse on woody herbaceous species in their earliest life stages (McGarvey et al. 2013), with overabundant herbivory negatively impacting seedling and sapling survival, growth and density (Found in McGarvey et al. 2013 – Dzieciolowski 1980, Gill and Beardall 2001, Healy 1997, Konig 1976, Putman et al. 1989).  Models predict that chronic overabundance will reduce understory diversity and decrease the abundance of traditionally dominant species (Holm et al. 2013).  Due to climate change and human activity, many non-endemic plant species are being introduced or number in forest ecosystems (Reference).  Deer find many of these species, including pawpaw (Asimina triloba), to be unpalatable (Found in McGarvey 2013 – Asnani 2006), and do not consume them at the same rates relative to their native counterparts (Reference), enabling them to form dense stands in forest understories (Found in Knauer et al. 2023 – Horsley and Marquis 1983; Stromayer and Warren 1997; Royo and Carson 2006).  This lack of browsing pressure, coupled with characteristics such as fast growth rate and greater adaptability to altered conditions precipitated by climate change, allow nuisance plant species to outcompete other species in the understory (Reference). The interaction of these nuisance species on the landscape contributes to a recruitment failure in forests, [*Where individuals from certain tree species do not enter the mature population - Do we need to define recruitment failure, or is that considered common knowledge?  If so, reference*).  The extent of the consequences derived from recruitment failure on species composition and forest structure is often not apparent for years (McGarvey et al. 2013)
